---
title: "Assignment2"
author: "Jose Rodriguez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
#Load Libraries
library(DBI)
library(RPostgres)
library(tcltk) #library to handle input prompts to hide password
sleep_default <- 3
```

##Prompt user for the database password

By using this method we make sure that our password remains confidential. The user is asked to type in the password when the code is executed. An alternative can be saving the password to a config file.

```{r Database password input}
#Create input prompt for database password
psswd <- .rs.askForPassword("Database Password:") #password is "postgres"
```

## Create the database connection between R and PostgresQL.

In this instance we connect to a locally hosted database. The same methodology may be used to connect to a cloud hosted database by using the proper login credentials and host link.


```{r create database connection}
#Create database connection
Sys.sleep(sleep_default)
con <- dbConnect(
  RPostgres::Postgres(),
  # without the previous and next lines, some functions fail with bigint data 
  #   so change int64 to integer
  bigint = "integer",  
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = psswd,
  dbname = "Data607_Assignment2")
```

# Check your connection

Once fully connected to the database it is good practice to check the tables at your disposal. The following code lists all the table names.

```{r Get list of tables in the database}
#List tables in database
dbListTables(con)
```


# Load table into a dataframe - Run query

Next we load the movies data into a dataframe by using a function built into DBI

```{r Load Data}
#Run query to load movies data from the database into a dataframe
df <- dbGetQuery(con, "select * from movies") #Returns the result of a query as a data frame
print(df) #view table
```

# Disconnect from Database

Lastly we terminate the connection to the database once we have the data we need.

```{r Disconnect from database}
#Disconnect from database
dbDisconnect(con)
```

