---
title: "Assignment7"
author: "Jose Rodriguez"
date: "2022-10-30"
output:
  pdf_document: default
  html_document: default
---

```{r, echo= FALSE}
#Set API key in a vector
api_key <- "GXydLE1KSp8UEvtev8gdeDh8ojwOVZIV"
```

# Overview

In this project we explore how to connect to an API. We specifically connect to the New York Times API to request the latest list of best-selling books. Ideally, we want to receive our data from the API in JSON format. Once stored locally in our machine, it'll be converted to a flat-file. In this case a dataframe. We want the dataframe to be "tidy", and therefore will execute some functions to unnest lists found in the json data.

In this project we use the following library packages:

* jsonlite
* httr
* tidyverse

```{r Load Libraries, include=FALSE}
library(jsonlite)
library(httr)
library(tidyverse)
```

# Connecting to API
In this section we set up our base url to make our initial request to the API. We make use of paste0() to seamlessly concatenate our private New York Times Developer API key. With the use of GET() we send our request and containerize it. Once we have our data locally, we check the status of the response by using status_code(). A 200 code means we successfully connected. Additionally, we use warn_for_status() to alert us of a failed connection; this function only executes if the connection is unsuccessful.


```{r pressure}
base_url <- "https://api.nytimes.com/svc/books/v3/lists/current/culture.json?api-key="
r <- GET(paste0(base_url, api_key)) #get name of best sellers list
status_code(r)

warn_for_status(r)
```
# Retrieving API Data

Next we use content() to retrieve the information stored in our variable. Second, we use fromJSON() to convert from a JSON object to an R object. Once an R object we can navigate the hierarchy until we reach our list of books [newdata\$results$books]. We combined the JSON to R conversion with as.data.frame() to get our data into a "flat" state.

However, our dataframe has 2 complex columns. The columns have dataframes as values.

```{r}
data <- content(r, as="text")
newdata <- fromJSON(data, flatten=TRUE)
books_df <- as.data.frame(newdata$results$books)

knitr::kable(head(books_df[, 1:26]), "simple")
```

# Tidying Data

The final task is to tidy our dataframe. In this case we chose to pivot wider, however, pivot longer can also be done. For sake of this project we pivot wider to reduce the number of observations. We pipe the unnest_wider function for each complex column. The result is a fully flat file.

```{r}
final_df <- books_df |>
  unnest_wider(isbns) |>
  unnest_wider(buy_links)

knitr::kable(head(final_df[, 1:28]), "simple")
```


