---
title: "Working with JSON, XML, and HTML Files"
author: "Jose Rodriguez"
date: "2022-10-17"
output:
  html_document: default
  pdf_document: default
---

```{r Load libraries, include=FALSE}
library(tidyverse)
library(jsonlite)
library(RJSONIO)
library(RCurl)
library(httr)
library(XML)
library(xml2)
library(rvest)
```

## Overview

In this document we explore way to load three different types of files into R; JSON, XML, and HTML. In addtion to loading the files we will tidy the information within these files to portray a flat file. In other words, we will transform them into data frames.

For each file type we first load the data from the github url. We then parse the information. Lastly, we transform it into a data frame class type.

```{r Table 1: JSON File}
url_json = "https://raw.githubusercontent.com/jlixander/DATA607/main/Assignment5/Assignment5.json"
json_df = jsonlite::fromJSON((url_json), simplifyDataFrame = TRUE) %>% 
  as.data.frame()

knitr::kable(json_df, "pipe")
```


```{r Table 2: XML File}
url_xml = GET("https://raw.githubusercontent.com/jlixander/DATA607/main/Assignment5/assignment5.xml") #retreive xml file
xml_doc <- xmlParse(url_xml) #parse xml file
xml_df <- xmlToDataFrame(xml_doc, nodes=getNodeSet(xml_doc, "//TopBooks")) #Select which level to unlist.

knitr::kable(xml_df, "pipe")
```


```{r Table 3: HTML File}
html_url <- "https://raw.githubusercontent.com/jlixander/DATA607/main/Assignment5/Assignment5.html"
html_doc <- xml2::read_html(html_url)
tables = html_doc |> html_table()
table_one = tables[[1]]
html_df <- head(table_one, - 1) #Remove extra blank row.

knitr::kable(html_df, "pipe")
```

# Differences
There are some distinguishable differences when looking at all three tables. On table one we can see that this parser created the column names by taking the full hierarchy of where the information is found. In a sense we can say that its given us the XML directory of the objects. Another difference is that the lists of authors were placed together in the same cell, delimited by a comma.If we take a closer look at our dataframe in our environment, we will find that we have data frames within a data frame.

On table 2 we have clean column names. One major difference is that the lists of authors were concatenated together in one single string.

On table 3 we also have clean column names. The difference in this table is that we have the list of author names delimited by pipes.